---
title: "Notropis hybrid analysis"
author: "K. Dye"
output: html_document
editor_options: 
  chunk_output_type: console
---


#NEW HYBRIDS-------------

#Load libraries and source files
```{r}
library(openxlsx)
```

#Read in popdata
```{r}
#retain cyt b individuals
popdata <- read.csv(file = "/home/kdye/Nmegalops/data/pop.data.csv", header = TRUE)

cytb <- popdata %>%
  filter(CYTB == "Y") 

cytb_inds <- cytb %>%
  select(INDV)

write.table(cytb_inds, "/home/kdye/Nmegalops/hybrid/newhybrids/cytb/cytb_inds",
col.names = FALSE, row.names = FALSE, quote = FALSE)
```

1) vcf file
  ## filter to keep only cytb individuals in vcf
```{bash}
cp -s ../../../genpops/notr_final.recode.vcf .

vcftools --vcf notr_final.recode.vcf --out notr_newhybrids --keep cytb_inds --recode --recode-INFO-all
```

2) File of individuals in group 1
```{r}
ama <- cytb %>%
  filter(SPECIES == "amabilis") %>%
  select(INDV)

write.table(ama, "/home/kdye/Nmegalops/hybrid/newhybrids/cytb/amabilis",
col.names = FALSE, row.names = FALSE, quote = FALSE)
```

3) File of individuals in group 2
```{r}
meg <- cytb %>%
  filter(SPECIES == "megalops") %>%
  select(INDV)

write.table(meg, "/home/kdye/Nmegalops/hybrid/newhybrids/cytb/megalops",
col.names = FALSE, row.names = FALSE, quote = FALSE)
```

#Preparing directories
```{bash}
cd ~/Nmegalops/hybrid/newhybrids/cytb
for i in $(seq 1 5); do mkdir run_$i; done
```

#Making files for new Hybrids
```{bash}
for i in $(seq 1 5); do 
echo "Processing run $i"
cd run_$i
Rscript ~/bin/NewHybrids_vcf_to_txt.r ~/Nmegalops/hybrid/newhybrids/cytb/notr_newhybrids.recode.vcf newhybrids_input.txt
cd ..
done
```

#Adding header
```{bash}
for i in $(seq 1 5); do 
INDV=$(vcfsamplenames ~/Nmegalops/hybrid/newhybrids/cytb/notr_newhybrids.recode.vcf | wc -l);
sed -i "1 s/^/NumIndivs $INDV\nNumLoci 150\nDigits 1\nFormat Lumped\n\n/" run_$i/newhybrids_input.txt;
done
```

#Adding the characterizations
```{bash}
nano genotype_frequencies.txt

Paste 
   5
 1_Bx      0.00000    0.250000    0.250000   0.50000
 0_Bx      0.50000    0.250000    0.250000   0.00000
 F1        0.00000     .5          .5        0.00000
 Pure_1    0.00000    0.00000     0.00000    1.00000
 Pure_0    1.00000    0.00000     0.00000    0.00000
```

#Running iterations
```{bash}
for i in $(seq 1 5); do
cd run_$i
RAND1=$(echo $((RANDOM)))
RAND2=$(echo $((RANDOM)))
echo $RAND1 $RAND2 > starting_seeds.txt
newhybs -d newhybrids_input.txt -c ../genotype_frequencies.txt -s $RAND1 $RAND2 --no-gui &
sleep 5
cd ..
done
}
```

#Formatting the output data
```{bash}
for i in $(seq 1 5); do
cd run_$i
vcfsamplenames ~/Nmegalops/newhybrids/cytb/notr_newhybrids.recode.vcf > tmp.indv
sed -i "1 s/^/Indv\n/" tmp.indv
paste tmp.indv aa-ScaledLikelihood.txt > aa-ScaledLikelihood_indv.txt
rm tmp.indv
Rscript ~/bin/NewHybrids_results_format.r amabilis ../amabilis megalops ../megalops
cd ..
done
```

```{R}
dat <- read.table("/home/kdye/Nmegalops/hybrid/newhybrids/cytb/run_1/aa-ScaledLikelihood_edit.txt", head=T, sep="\t")
dat$Max <- apply(dat[,2:ncol(dat)],1,max)
wb <- write.xlsx(dat, "MxA.xlsx", colNames = T, sheetName="run1")
dat2 <- read.table("/home/kdye/Nmegalops/hybrid/newhybrids/cytb/run_2/aa-ScaledLikelihood_edit.txt", head=T, sep="\t")
dat2$Max <- apply(dat2[,2:ncol(dat2)],1,max)
addWorksheet(wb, sheetName = "run2")
writeData(wb, "run2", dat2)
dat3 <- read.table("/home/kdye/Nmegalops/hybrid/newhybrids/cytb/run_3/aa-ScaledLikelihood_edit.txt", head=T, sep="\t")
dat3$Max <- apply(dat3[,2:ncol(dat3)],1,max)
addWorksheet(wb, sheetName = "run3")
writeData(wb, "run3", dat3)
dat4 <- read.table("/home/kdye/Nmegalops/hybrid/newhybrids/cytb/run_4/aa-ScaledLikelihood_edit.txt", head=T, sep="\t")
dat4$Max <- apply(dat4[,2:ncol(dat4)],1,max)
addWorksheet(wb, sheetName = "run4")
writeData(wb, "run4", dat4)
dat5 <- read.table("/home/kdye/Nmegalops/hybrid/newhybrids/cytb/run_5/aa-ScaledLikelihood_edit.txt", head=T, sep="\t")
dat5$Max <- apply(dat5[,2:ncol(dat5)],1,max)
addWorksheet(wb, sheetName = "run5")
writeData(wb, "run5", dat5)
addWorksheet(wb, sheetName = "Summary")
saveWorkbook(wb, "/home/kdye/Nmegalops/hybrid/newhybrids/cytb/MxA.xlsx", overwrite = TRUE)
```


#ADEGENET - HYBRIDIZE-------------

#Adegenet Hybrid simulations

#Load libraries and source files
```{r}
library(adegenet)
library(vcfR)
library(reshape2)
library(ggplot2)
library(ggforce)

source("/home/kdye/bin/genind.R")
source("/home/kdye/bin/PCA.R")
source("/home/kdye/bin/libraries.R")
source("/home/kdye/bin/ggplot.R")
source("/home/kdye/bin/VCFfilterstats.R")
```

#Input files

1) vcf file
```{bash}
cp -s ../../../genpops/notr_final.recode.vcf .
```

#Read in popdata
  ##filter to retain cyt b individuals
```{r}
popdata <- read.csv(file = "/home/kdye/Nmegalops/pop.data.csv", header = TRUE)

cytb <- popdata %>%
  filter(CYTB == "Y")

non_cytb <- popdata %>%
  filter(CYTB == "N") %>%
  select(INDV)
```

2) File of individuals in group 1
```{r}
ama <- popdata %>%
  filter(SPECIES == "amabilis") %>%
  select(INDV)

write.table(ama, "/home/kdye/Nmegalops/adegenet/cytb/amabilis",
col.names = FALSE, row.names = FALSE, quote = FALSE)
```

3) File of individuals in group 2
```{r}
meg <- popdata %>%
  filter(SPECIES == "megalops") %>%
  select(INDV)

write.table(meg, "/home/kdye/Nmegalops/adegenet/cytb/megalops",
col.names = FALSE, row.names = FALSE, quote = FALSE)
```

#Making the input files
```{bash}
cd ~/Nmegalopsadegenet/cytb

#Adding a header
echo -e "Sample_ID\tSpecies"> species.txt
#Adding species 1 individuals and the species
awk -v OFS="\t" '{print $0, "amabilis"}' ~/Nmegalops/adegenet/cytb/amabilis >> species.txt
#Adding species 2 individuals and the species
awk -v OFS="\t" '{print $0, "megalops"}' ~/Nmegalops/adegenet/cytb/megalops >> species.txt
```

#Importing data
```{r}
vcf <- read.vcfR(file = "~/Nmegalops/adegenet/cytb/notr_final.recode.vcf")
gen.vcf <- vcfR2genind(vcf)

#Assign strata
strata(gen.vcf) <- popdata[match(indNames(gen.vcf),popdata$INDV),]
setPop(gen.vcf) <- ~SPECIES

remove_non_cytb <- non_cytb$INDV
cytb.gen <- gen.ind.rem.Ind(gen.vcf, remove_non_cytb)

#Re-assign strata
strata <- read.table(file = "~/Nmegalops/adegenet/cytb/species.txt", header = TRUE, sep="\t")
strata(cytb.gen) <- strata[match(indNames(cytb.gen),strata$Sample_ID),]
head(cytb.gen@strata)

rm(vcf)

#Define Strata
setPop(cytb.gen) <- ~Species
temp <- seppop(cytb.gen)
ama <- temp$`amabilis`
meg <- temp$`megalops`

#Simulate samples
F1 <- hybridize(ama, meg, n = 100, pop = "amaxmeg")
ama_bx <- hybridize(ama, F1, n = 100, pop = "ama_bx")
meg_bx <- hybridize(meg, F1, n = 100, pop = "meg_bx")
F2 <- hybridize(F1, F1, n= 100, pop = "F1XF1")
pooled_gens <- repool(gen.vcf, F1, ama_bx, meg_bx, F2)

save(pooled_gens, file="~/Nmegalops/adegenet/cytb/ama_x_meg_sims.gz", compress=T)
#pooled_gens <- load("~/Nmegalops/adegenet/cytb/ama_x_meg_sims.gz")
```

#Visualize results

#PCA

```{r}
#Run PCA
x <- scaleGen(pooled_gens, NA.method = "mean")
pca <- dudi.pca(x,center = FALSE, scale = FALSE, scannf = FALSE, nf = 4)
eig <- eigenvalues(pca)
```

#Plot
```{r}
# Define colors
map_colors <- c("#007eb0", "#b53333")

col <- c("#007eb0", "#b53333", 'darkorchid3', "darkorange3",'#799c41', '#000000')

# Assuming pca$li contains the principal component scores
# Assuming pop(pooled_gens) contains the group information

# Calculate means and standard deviations for each group along Axis1 and Axis2
group_means <- aggregate(pca$li[, c("Axis1", "Axis2")], by = list(pop(pooled_gens)), FUN = mean)
group_sds <- aggregate(pca$li[, c("Axis1", "Axis2")], by = list(pop(pooled_gens)), FUN = sd)

# Set the desired confidence level for the ellipses
confidence_level <- 0.95

# Calculate the radius of the circles based on the standard deviations
group_sds$radius_dim1 <- group_sds$Axis1 * qnorm(confidence_level)
group_sds$radius_dim2 <- group_sds$Axis2 * qnorm(confidence_level)

# Create a data frame for plotting the circles
circle_data <- merge(group_means, group_sds, by = "Group.1")
circle_data <- circle_data[, c("Group.1", "Axis1.x", "Axis2.x", "radius_dim1", "radius_dim2")]
colnames(circle_data) <- c("group", "center_axis1", "center_axis2", "radius_dim1", "radius_dim2")

#remove elipsis from amabilis and megalops individuals
circle_data_subset <- circle_data[-c(2,6), ]

# Reorder levels in the dataframe
pca$li$pop_pooled_gens <- factor(pop(pooled_gens), levels = c("megalops", "amabilis", "amaxmeg", "ama_bx", "meg_bx", "F1XF1"))
#pca$li$pop_pooled_gens <- factor(pca$li$pop_pooled_gens, levels = c("megalops", "amabilis", "amaxmeg", "ama_bx", "meg_bx", "F1XF1"))

#Kayla's ggplot  
#created a subset to remove the ellipse from the data points.. 
ggplot() +
  geom_point(data = pca$li[1:length(indNames(gen.vcf)),], 
             aes(x = Axis1, y = Axis2, fill = as.factor(pop_pooled_gens[1:length(indNames(gen.vcf))])),
             size = 2, shape = 21) +
    labs(x = paste("Axis1:", round(eig[1, 3], digits = 3), "%"),
       y = paste("Axis2:", round(eig[2, 3], digits = 3), "%")) +
  scale_fill_manual(values = map_colors, 
                    limits = c("megalops", "amabilis")) +
  geom_ellipse(data = circle_data_subset, aes(x0 = center_axis1, y0 = center_axis2, 
                                       a = radius_dim1*4, b = radius_dim2*4, 
                                       angle = 0, 
                                       group = group, color = group), size = 0.3) +
                 scale_color_manual(values = col[3:6], 
                     limits = c("amaxmeg", "ama_bx", "meg_bx", "F1XF1")) +
  theme_bw() + 
  labs(fill = "Species", color = "Simulation")

